# ============================================================================
# RECOVERED HISTORICAL RESEARCH SCRIPT
# Original relative path: SYNTOR-NO-SI\climate data analysis.Rmd
# Source SHA256: 88C873035DC51F156AF5D099381752AF8499E24B3AE96FA6E5AE0A6C0765F24D
#
# This is a sanitized copy recovered from the historical WEPP research workspace.
# Machine-specific absolute paths and email addresses were redacted.
# The scientific statements/calculations were otherwise retained as historical
# evidence. This file is not guaranteed to run without reconstructing its
# historical data objects, package versions, and upstream workflow state.
# ============================================================================
---
title: "Precipitation Analysis"
author: "Lifeng Yuan"
date: "Nov. 2020"
output: html_document
contact: <EMAIL_REDACTED>
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R script was developed to analyze some basic elements for climate change,such as precipitation, maximum and minimum temperature by monthly or yearly.It supports the WEPP project at USDA-ARS Grazinglands Research Center, El Reno,Oklahoma.                                               
# Setting the working directory and load working packages
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
getwd()
setwd("<LOCAL_PATH_REDACTED>")
library(readxl)
library(openxlsx)
library(tidyverse)
library(hydroTSM)
library(zoo)
```

## This block code is conducted to read in the excel file

Get the name of each sheet in a excel file, read each sheet in a entire excel file,names the entire workbook by different names of sheets, and create a new variable with a simple name to store the entire.workbook

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
sheets_name <- openxlsx::getSheetNames('<LOCAL_PATH_REDACTED>')
entire.workbook <- lapply(excel_sheets('Future2_RCP45.xlsx'),function (x) read_excel(path='Future2_RCP45.xlsx',sheet = x,skip=14))
names(entire.workbook) <- sheets_name
cli <- entire.workbook
cli
```
Only read in precipitation, maximum and minimum temperature in the 'cli' list
```{r}
for (i in 1:length(cli)){
  cli[[i]] <- as.data.frame(cli[[i]])
  cli[[i]] <- cli[[i]][,c(1,2,3,4,8,9)]  
  colnames(cli[[i]]) <- c('day','mon','year','prcp','tmax','tmin')
}
```

## Summary monthly and yearly precipitation,and write them into 'cli' data frame
Note:
monthly precipitation amount were calculated by summarize all daily precipitation in that month;
yearly precipitation amount were calculated by summarize all monthly precipitation in that year;
average monthly maximum temperature were calculated by adding daily maximum temperature for a month then dividing the number of days in a month;
average yearly maximum temperature were calculated by adding monthly maximum temperature for a year then dividing 12 months;
average monthly minimum temperature were calculated by adding daily minimum temperature for a month then dividing the number of days in a month;
average yearly minimum temperature were calculated by adding monthly minimum temperature for a year then dividing 12 months.
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
for (i in 1:length(cli)){
  cli[[i]] %>% group_by(mon,year) %>%
    mutate(prcp_mon = sum(prcp)) -> cli[[i]]
  cli[[i]] %>% group_by(year) %>%
    mutate(prcp_year = sum(prcp)) -> cli[[i]]
  cli[[i]] %>% group_by(mon,year) %>%
    mutate(avg_tmax_mon = sum(tmax)/n()) -> cli[[i]]
  cli[[i]] %>% group_by(year) %>%
    mutate(avg_tmax_year = sum(tmax)/n()) -> cli[[i]]
  cli[[i]] %>% group_by(mon,year) %>%
    mutate(avg_tmin_mon = sum(tmin)/n()) -> cli[[i]]
  cli[[i]] %>% group_by(year) %>%
    mutate(avg_tmin_year = sum(tmin)/n()) -> cli[[i]]
}
```
# Build a null excel workbook and write all summary with the corresponding sheet names
```{r echo=TRUE}
wb <- createWorkbook()
for (i in 1:26) {
  addWorksheet(wb,sheets_name[i])
  writeDataTable(wb,i,cli[[i]])
}
```
```{r echo=FALSE}
saveWorkbook(wb,file="./Future2_RCP45_sum.xlsx",overwrite = TRUE)
```
Showing a summary of GPCC_baseline for an example
```{r}
summary(cli[[1]])
```

Adding a date column into the 'cli' for further analysis. I assume that the start date is from 2005-01-01 for the present climate baseline although its first day could be any day during 2005-01-01 to 2019-01-01. 

```{r}
for (i in 1:26){
  print(paste(names(cli[i]),"sheet has: ",nrow(cli[[i]]),' rows'))
  if (i == 1){
    myDate <- data.frame(date = seq(as.Date('2005/1/1'),as.Date('2104/12/7'),'days'))
    cli[[1]] <- data.frame(myDate,cli[[1]])
  } else {
    myDate <- data.frame(date = seq(as.Date('2021/1/1'),as.Date('2120/12/7'),'days'))
    cli[[i]] <- data.frame(myDate,cli[[i]])
  }
}
```
# Precipitation and temperature analysis
Building a list only including 'date' and 'precipitation'
```{r include=FALSE}
myList <- list()
for (i in 1:26){
  myList[[i]] <- cli[[i]][,c('date','prcp')]
}
names(myList) <- sheets_name
names(myList)
```
Plotting monthly and annual hydrology 
```{r include=FALSE}
for (i in 1:26){
  z <- read.zoo(myList[[i]])
  m <- daily2monthly(z,FUN = sum)
  # x <- window(z,start = as.Date('2055-01-01'))
  hydroplot(m,var.type = 'Precipitation', main=sheets_name[i],
            pfreq = 'ma')
}
```
## Showing monthly and annual time series precipitation of GPCC_baseline for an example
```{r}
hydroplot(m,var.type = 'Precipitation', main=sheets_name[1],
            pfreq = 'ma')
```
Building a list including 'date','precipitation','maximum temperature', and 'minimum temperature'
```{r include=FALSE}
cli1 <- list()
for (i in 1:26){
  cli1[[i]] <- cli[[i]][,c('date','prcp','tmax','tmin')]
  pcp <- read.zoo(cli1[[i]][,c(1,2)])
  tmx <- read.zoo(cli1[[i]][,c(1,3)])
  tmn <- read.zoo(cli1[[i]][,c(1,4)])
  m <- climograph(pcp=pcp,tmx=tmx,tmn=tmn,
                  na.rm = TRUE,main = paste('Climograph at',sheets_name[i]))
}
```
Showing the first 50 rows in a cli1[[15]]
```{r}
head(cli[[15]],50)
```

# Showing one climograph for an example
```{r}

pcp <- read.zoo(cli1[[26]][,c(1,2)])
tmx <- read.zoo(cli1[[26]][,c(1,3)])
tmn <- read.zoo(cli1[[26]][,c(1,4)])
m <- climograph(pcp=pcp,tmx=tmx,tmn=tmn,na.rm = TRUE,main =   paste('Climograph at',sheets_name[26]))
```
# Plotting monthly precipitation for years
```{r}
z <- read.zoo(cli[[2]][,c(1,2)])
m <- daily2monthly(z,FUN = sum)
M <- matrix(m, ncol=12, byrow=TRUE)
colnames(M) <- month.abb
rownames(M) <- unique(format(time(m),'%Y'))
require(lattice)
print(matrixplot(M, ColorRamp = 'Precipitation',
                 main = 'Monthly precipitation at WEATHERFORD OK,[mm/month]'))
```

